---
title: "Data Normalisation"
author: "Justin Lo"
date: "`28_July_2023"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
This is the dataset needed to be normalised

  '| EmployeeID | SalesPerson | SalesOffice | OfficeNumber | Customer1 | Customer2 | Customer3 |
   | ------:|:-------------:|:----------:|:------------ |:-------:|:------:|:------:|
   | 1003   | Mary Smith    |  Chicago   | 312-555-1212 | Ford   | GM |       |
   | 1004   | John Hunt     |  New York  | 212-555-1212 | Dell   | HP | Apple |
   | 1005   | Martin Hap    |  Chicago   | 312-555-1212 | Boeing |    |       |'
```{r}
library('janitor')
library('tidyverse')

rmd.string <-
  '| EmployeeID | SalesPerson | SalesOffice | OfficeNumber | Customer1 | Customer2 | Customer3 |
   | ------:|:-------------:|:----------:|:------------ |:-------:|:------:|:------:|
   | 1003   | Mary Smith    |  Chicago   | 312-555-1212 | Ford   | GM |       |
   | 1004   | John Hunt     |  New York  | 212-555-1212 | Dell   | HP | Apple |
   | 1005   | Martin Hap    |  Chicago   | 312-555-1212 | Boeing |    |       |'

rmd.table <- read_delim(rmd.string, delim = '|')[-1,2:8]
#To read the hand-typed dataset and creating a dataframe fo rit

colnames(rmd.table) <- str_squish(colnames(rmd.table))

rmd.table <- 
  rmd.table %>% 
  mutate(EmployeeID = as.integer(EmployeeID),
         across(SalesPerson:Customer3, ~ str_squish(.x)))
#set employeeID as integer and clear white spaces in columns.


rmd.table<- rmd.table %>% 
  pivot_longer(cols = starts_with('Customer'),
               values_to = 'Customer') %>% 
  filter(Customer != '')
#pivot the table into rectangular shape and only take the value where customer is not blank 
```
1NF achieved

```{r}
employee<- rmd.table %>% 
  distinct(EmployeeID, SalesPerson, SalesOffice, OfficeNumber)

customer<- rmd.table %>%  
  select(EmployeeID, Customer)
#creating employee and customer table 
```
Now all data depend on their primary key and that there is no partial dependency, 2NF achieved

```{r}
office<- employee %>% 
  select(OfficeNumber, SalesOffice)

employee<- employee %>% 
  select(EmployeeID, SalesPerson, OfficeNumber)
#creating pffice and employee table
```
New table for offices as the primary key does not fully define SalesOffice(it can also be defined by OfficeNumber)
Now that there is no transitive partial dependency, 3NF achieved 

Another dataset to be normalised

| Name           | Address    |   Movies Rented   |  Salutation  | Category | Fiction? |
   |:---------------|:-----------|:------------------|:------------:|----------|----------|
   | Bob Smith      | 1 Houghton Street  | _Star Wars_, _Fast and Furious 7_ |  Dr. |  Scifi, Action | Yes, Yes|
   | Pry Ministair  | 10 Downing St  |  _Brexit: Three Years On_   |  Rt Hon.  | Documentary | No |
   | Joe Bloggs     | 8 Myhatt St.  |  _The Big Short_, _Star Wars_  | Mr. | Documentary, Scifi | No, Yes |
   
```{r}
movie.string<-  '| Name           | Address    |   Movies Rented   |  Salutation  | Category | Fiction? |
   |:---------------|:-----------|:------------------|:------------:|----------|----------|
   | Bob Smith      | 1 Houghton Street  | _Star Wars_, _Fast and Furious 7_ |  Dr. |  Scifi, Action | Yes, Yes|
   | Pry Ministair  | 10 Downing St  |  _Brexit: Three Years On_   |  Rt Hon.  | Documentary | No |
   | Joe Bloggs     | 8 Myhatt St.  |  _The Big Short_, _Star Wars_  | Mr. | Documentary, Scifi | No, Yes |'
movie.table<-read_delim(movie.string)[-1,2:7]

movie.table<- movie.table %>% 
  mutate(across(everything(), ~ str_squish(str_remove_all(.x, '\\_')))) %>% 
  janitor::clean_names() %>% 
  separate_rows(c(movies_rented, category, fiction), sep = ',') 
#clearing white spaces, tidying columns name. Splitting the multiple values in the same cell into different rows
```
1NF achieved 

There is partial dependency
```{r}
movie.table$userID<-c(001,001,002,003,003)
movie.table$moviesID<-as.character(c(1,2,3,4,1))

#rearranging columns in movie.table
movie.table<- movie.table %>% 
  select(userID, name, address, moviesID,movies_rented, salutation, category, fiction)

#create rental dataset
rentals<- movie.table %>% 
  distinct(userID, moviesID)

#create movies dataset
movies<- movie.table %>% 
  distinct(moviesID, movies_rented, category, fiction) %>% 
  mutate(across(everything(), ~ str_squish(.x)))
movies<-movies[-5,]

#create user dataset 
user<- movie.table %>% 
  distinct(userID, name, address, salutation)
```
2NF achieved 

As there is transitive dependency between category and fiction in the movie table, we have to make a new table and update the movie table.
```{r}
movies$categoryID<-as.character(c(1,2,3,3))
movies<- movies %>% 
  select(moviesID,movies_rented,categoryID,category,fiction)

#create category table
category<- movies %>% 
  distinct(categoryID,category,fiction)
  
#update movies table
movies<- movies %>% 
  select(moviesID, movies_rented, categoryID)
```
3NF achieved

To save the two tables into csv file
```{r}
write.csv(rmd.table, file = 'sales.csv', row.names = FALSE)
write.csv(movie.table, file = 'movie.csv', row.names = FALSE)
```


   